---
title: "pe_project - models"
author: "laispetri"
date: "`r format(Sys.time(), "%m/%d/%Y")`"
output: html_document
editor_options: 
  chunk_output_type: console
---

**Please, read these notes before running the code:**

1. You will need to create the following folders in the working directory: "data", "figures", and "modelparameters".
2. Paste the files from this webpage (https://doi.org/10.5061/dryad.mpg4f4r5w) inside of the "data" folder you just created.
3. Note: figures included in the main manuscript or SM will not be displayed in the html file after knitting. They will be inside of the folder you created, "figures". The aspect ratio could not be maintained when knitting (probably due to my own coding limitation).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r packages and directory, echo=FALSE}
# loading the packages
## data wrangling
require(tidyverse) 
## models
require(rjags)
require(coda)
require(broom)
## color blind discrete colors
require(pals) 
```

Paths to import files, and export outputs
```{r}
workingdir <- getwd()
datain <- file.path(workingdir, "data")
graph.wd <- file.path(workingdir, "figures")
dataout <- file.path(workingdir, "modelparameters")
```

Defining colors
```{r}
sitecolor<-as.vector(alphabet(4))
sitecolor2<-c("#F0A0FF","#cce3f8")
mocolor<-as.vector(alphabet2(3))
mocolor2 <- c("#DD9EFE","#AA0DFE","#ADCDFE","#3283FE","#CEC19E","#85660D")
```

# Cover time-series models
## Importing data sets:
```{r datasets}
data <- as_tibble(read.csv(file.path(datain,'data.csv'),
                              header = T, stringsAsFactors = F))

richnessN <- as_tibble(read.csv(file.path(datain,'richnessN.csv'), 
                            header = T, stringsAsFactors = F))

coverI2019 <- as_tibble(read.csv(file.path(datain,'coverI2019.csv'), 
                                 header = T, stringsAsFactors = F)) 

dominantI <- as_tibble(read.csv(file.path(datain,'dominantI.csv'), 
                                  header = T, stringsAsFactors = F))

traitsSeedMix <- as_tibble(read.csv(file.path(datain,'traitsSeedMix.csv'), 
                                  header = T, stringsAsFactors = F)) 

traitsDiversity <- as_tibble(read.csv(file.path(datain,'traitsDiversity.csv'), 
                                  header = T, stringsAsFactors = F))

VPD <- read.csv(file.path(datain,"VPD.csv"), header = TRUE, stringsAsFactors = FALSE)
``` 

## Model
Organizing variables
```{r data formatting model}
datamodel <- list (P=length(unique(data$ID)), 
                   S=length(unique(data$subplot)),
                   treatment=data$subplot,
                   Y=3, #number of years
                   Cover=as.matrix(data[,6:8]),
                   RichN=as.matrix(data[,9:11]),
                   sm=as.matrix(data[,12:14]),
                   Cover2019=data$Cover2019,
                   grassAG20=data$grassAG1,
                   forbsAG20=data$forbAG1,
                   woodyAG20=data$woodyAG1,
                   grassAG21=data$grassAG2,
                   forbsAG21=data$forbAG2,
                   woodyAG21=data$woodyAG2,
                   light=as.matrix(data[,24:26]),
                   SLA=as.matrix(data[,21:23]),
                   ID=data$ID)
```

Initials
```{r}
inits <- list(
  list(B = matrix(c(rep(-30, times=3*length(unique(data$subplot)))),ncol=3),
       tau = c(rep(1, times = 1)),
       alpha = c(rep(-1, times=4)),
       w = c(rep(0.33, times=3))
       ),
  list(B = matrix(c(rep(0, times=3*length(unique(data$subplot)))),ncol=3),
       tau = c(rep(0.1, times = 1)),
       alpha = c(rep(0, times=4)),
       w = c(rep(0.33, times=3))
       ),
  list(B = matrix(c(rep(30, times=3*length(unique(data$subplot)))),ncol=3),
       tau = c(rep(0.5, times = 1)),
       alpha = c(rep(1, times=4)),
       w = c(rep(0.33, times=3))
       )
)
```

Model code
```{r FinalModel}
FinalModel<-"model{

 for(i in 1:P){ #number of subplots
     for(t in 1:Y){ #number of years
        
        Cover[i,t]~dnorm(C[i,t],tau) #likelihood
        Cover.pred[i,t]~dnorm(C[i,t],tau) #predicted values
        
        residuals[i,t]<-Cover.pred[i,t]-Cover[i,t] #calculating residuals
        
      }
        
    C[i,1]<-B[treatment[i],1]*Cover2019[i]+alpha[1]*RichN[i,1]+alpha[2]*SLA[i,1]+alpha[3]*light[i,1]+alpha[4]*sm[i,1] #process model for 2020
     
    C[i,2]<-B[treatment[i],2]*(w[1]*grassAG20[i]+w[2]*forbsAG20[i]+w[3]*woodyAG20[i])+alpha[1]*RichN[i,2]+alpha[2]*SLA[i,2]+alpha[3]*light[i,2]+alpha[4]*sm[i,2] #process model for 2021
     
    C[i,3]<-B[treatment[i],3]*(w[1]*grassAG21[i]+w[2]*forbsAG21[i]+w[3]*woodyAG21[i])+alpha[1]*RichN[i,3]+alpha[2]*SLA[i,3]+alpha[3]*light[i,3]+alpha[4]*sm[i,3] #process model for 2022
    

  }

# priors
for(i in 1:4){
alpha[i]~dnorm(0,0.01) 
}

for(i in 1:S){ #number of treatments
for(t in 1:Y){
    B[i,t]~dnorm(0,0.1)
}
}

w[1:3]~ddirich(W[])
for(i in 1:3){
W[i]<-1
}

tau~dgamma(0.0001,0.0001)
varC<-1/tau

} #end of model"

#exports model
write.table(FinalModel,file.path(dataout,"FinalModel.txt"), sep = "\t",
            row.names = TRUE, col.names = NA)

```

Compile
```{r}
m <-jags.model(textConnection(FinalModel), datamodel, inits, n.chains=3)
```

Burn-in
```{r}
update(m, 10000)
```

Running model and retrieving parameters
```{r}
parameters<-coda.samples(m, c("B", "tau", "varC", "alpha", "w"), 
                         n.iter = 50000, thin = 100)
parameters0<-summary(parameters)
```

Plot: Posterior distributions
```{r} 
parametersB<-coda.samples(m, c("B"), n.iter = 50000, thin = 100)
pdf(paste0(file.path(dataout,"PosteriorDistributionsB.pdf")))
par(mar = c(2, 2, 2, 2)) 
plot(parametersB)
dev.off()

pdf(paste0(file.path(dataout,"PosteriorDistributionsLikelihoodTau.pdf")))
plot(parameters[,'tau'], main="tau")
plot(parameters[,'varC'], main="varC")
dev.off()

parametersalpha<-coda.samples(m, c("alpha"), n.iter = 50000, thin = 100)
pdf(paste0(file.path(dataout,"PosteriorDistributionsAlpha.pdf")))
par(mar = c(2, 2, 2, 2)) 
plot(parametersalpha)
dev.off()

parametersW<-coda.samples(m, c("w"), n.iter = 50000, thin = 100)
pdf(paste0(file.path(dataout,"PosteriorDistributionsW.pdf")))
par(mar = c(2, 2, 2, 2)) 
plot(parametersW)
dev.off()
```

Exporting parameter values, predictions and residuals
```{r}
Quantiles <- as.data.frame(parameters0[2]) %>% mutate_if(is.numeric, round,digits=4) %>% rownames_to_column(., var = "ParameterID") %>% select(1,2,6)
Stats <- as.data.frame(parameters0[1]) %>% mutate_if(is.numeric, round,digits=4) %>% rownames_to_column(., var = "ParameterID") %>% select(1:3) %>% left_join(Quantiles)
write_csv(Stats, file.path(dataout,"ParameterValues.csv"))

predicted <- coda.samples(m, c("Cover.pred"), n.iter=10000, thin = 1)
 pred <- as.data.frame(summary(predicted)$statistics)
pred <- pred %>% rename_all(function(x) paste0(x,"_pred")) %>% rownames_to_column(., var = "modelID") %>% mutate(ID=row_number())
write_csv(pred, file.path(dataout,"Predicted.csv"))

res <- coda.samples(m, c("residuals"), n.iter=10000, thin = 1)
res <- as.data.frame(summary(res)$statistics)
res <- res %>% rename_all(function(x) paste0(x,"_res")) %>% rownames_to_column(., var = "modelID_res") %>% mutate(ID=row_number())
write_csv(res, file.path(dataout,"Residuals.csv"))

```

Merging model predictions and raw data
```{r}
predModel <- data %>% pivot_longer(c(Cover1,Cover2,Cover3)) %>% group_by(name,ID) %>% mutate(ID=cur_group_id()) %>% ungroup() %>% left_join(res) %>% rename(residuals=Mean_res) %>% left_join(pred) %>% mutate(site = str_sub(forest_stand,end=-2), year=ifelse(name=="Cover1",2020,ifelse(name=="Cover2",2021,2022))) %>% left_join(dominantI) %>% 
  left_join(coverI2019) %>% distinct() %>% mutate(sppName=ifelse(is.na(sppName), "No invader",sppName))
glimpse(predModel)
```

Plot: predicted against observed
```{r fit}
fit <- lm(value ~ Mean_pred - 1, data = predModel)
ggplot(data = predModel, aes(y = value, x = Mean_pred)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  stat_function(fun = function(x) predict(fit, newdata = data.frame(Mean_pred = x)),
                color = "black", linewidth = 2) +
  annotate(label = sprintf("y = %.3f x\nR² = %.2f", coef(fit), summary(fit)$r.squared),
           geom = "text", x = 10, y = 90, size = 5) +
  geom_point(shape=19,size=2,alpha=0.5) +
  scale_y_continuous(limits=c(0,100))+
  scale_x_continuous(limits=c(0,NA))+
  ylab("Observed values")+
  xlab("Predicted values")+ 
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 17),
        strip.text.x = element_text(size = 15),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      plot.title = element_text(hjust = 0.5,size=18,color="darkblue"),
      panel.spacing = unit(1, "lines"))

ggsave(file.path(graph.wd,"ModelFit.png"), plot = last_plot(), width=7, height=5, dpi = 600, units="in")
    
```

## Results
Wrangling data to plot hyper parameter B
```{r}
BETAmodelpar<- as.data.frame(parameters0[2]) %>% 
  rownames_to_column(., var = "parameter") %>% 
  filter(grepl("^B",parameter)) %>% mutate(ID=rep(1:6,3),
                                            year=rep(2020:2022,times=1,each=6),
                                            subplot=rep(1:6,times=3),
                                            filling=ifelse(`quantiles.2.5.`<0&`quantiles.97.5.`<0|
                                                              `quantiles.2.5.`>0&`quantiles.97.5.`>0,
                                                            "black","white")) %>% 
  left_join(distinct(dplyr::select(data,c("subplot","subplotOriginal")))) %>% distinct() %>% 
  rename(subplotNEW=subplot,
         subplot=subplotOriginal)
```

(Q1) Could seed additions be a successful practice driving community assembly after invasive plant removal?
```{r}
DataPlot <- BETAmodelpar %>% 
  filter(subplot==1|subplot==7|subplot==4|subplot==10|subplot==3|subplot==9) %>%
  arrange(subplot) %>% 
  mutate(Treatment=ifelse(subplot<7,1,2),
         Treatment=ifelse(year==2021,Treatment+2,
                          ifelse(year==2022,Treatment+4,Treatment))) %>% 
  arrange(year) %>% 
  mutate(guide = rep(1:3, times=6)) %>% 
  arrange(year,guide) %>% 
  mutate(index=seq(1,length=18,by=0.3),
         index=ifelse(year==2021,index+0.2,
                      ifelse(year==2022,index+0.4,index)),
         type=ifelse(subplot==3|subplot==9,"Forbs and\ngrasses added",ifelse(subplot==4|subplot==10,"Forbs added","Control")),
         removal=ifelse(subplot<5,"One-time removal","Multi-year removal"))

names(mocolor2) <- c("1","2","3","4","5","6")

DataPlot$type <- factor(DataPlot$type, levels = c("Control","Forbs and\ngrasses added","Forbs added"))

DataPlot %>% 
  ggplot(aes(y=`quantiles.50.`,x=index,color=as.factor(Treatment),shape=type))+  
  geom_errorbar(aes(ymin=`quantiles.2.5.`, ymax=`quantiles.97.5.`),
                width=.1, linewidth=1.3)+
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(size=6,aes(shape=as.factor(type)),stroke=2)+
  scale_y_continuous(n.breaks=6)+
  scale_x_continuous(breaks=c(1.85,3.85,5.85),
                     labels=c("2020","2021","2022"),
                     expand=c(0,0.3))+
  scale_color_manual(values=mocolor2, 
                     breaks=c("1","2"),
                     labels=c("One-time removal","Multi-year removal"),
                     guide=guide_legend(override.aes = list(shape=16,size=4,linetype = c(0,0),
                                                            color="black",alpha=c(0.2,1))))+
  ylab("Native community recovery rate\n(mean+95%CI)")+
  theme_classic()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.text.x = element_text(size=22,color = c(mocolor2[2], mocolor2[4], mocolor2[6])),
        axis.title = element_text(size = 20, face="bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(vjust = 1),
        legend.margin=margin(-4,3,2,1),
        legend.position = c(0.17,0.74),
        legend.text = element_text(size=15),
        legend.background = element_blank(),
        legend.title = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.spacing = unit(1, "lines")) +   
   guides(shape = guide_legend(override.aes = list(shape=c(1,2,0),size=3.5)))+
  # annotate("text", x=0.9,y=3.6,label="(A)",fontface="bold",size=6) +
  annotate("text", x=1,y=0.75,label="ac",size=5) +
  annotate("text", x=1.3,y=0.87,label="ac",size=5) +
  annotate("text", x=1.6,y=0.70,label="a",size=5) +
  annotate("text", x=1.9,y=0.90,label="ac",size=5) +
  annotate("text", x=2.2,y=0.58,label="a",size=5) +
  annotate("text", x=2.5,y=0.56,label="a",size=5) +
  annotate("text", x=3,y=1.40,label="be",size=5) +
  annotate("text", x=3.3,y=2.05,label="be",size=5) +
  annotate("text", x=3.6,y=1.22,label="bce",size=5) +
  annotate("text", x=3.9,y=1.99,label="be",size=5) +
  annotate("text", x=4.2,y=1.05,label="bcd",size=5) +
  annotate("text", x=4.5,y=1.97,label="be",size=5) +
  annotate("text", x=5,y=0.87,label="bcd",size=5) +
  annotate("text", x=5.3,y=1.15,label="bcd",size=5) +
  annotate("text", x=5.6,y=0.70,label="bcd",size=5) +
  annotate("text", x=5.9,y=1.09,label="bcd",size=5) +
  annotate("text", x=6.2,y=0.53,label="bcd",size=5) +
  annotate("text", x=6.5,y=1.08,label="bcd",size=5)

ggsave(file.path(graph.wd,"Q1.png"), plot = last_plot(), width=8, height=5, dpi = 600, units="in")
```

Contribution of growth forms to treatment effects over time
```{r}
GFmodelpar <- as.data.frame(parameters0[2]) %>% 
  rownames_to_column(., var = "parameter") %>% 
  filter(grepl("^w",parameter)) 

GFmodelpar %>%  
  ggplot(aes(y=`quantiles.50.`,x=parameter))+
  geom_errorbar(aes(ymin=`quantiles.2.5.`, ymax=`quantiles.97.5.`),width=.1, linewidth=1.3, color="black")+
  geom_point(shape=18,size=9) + 
  scale_x_discrete(labels=c("Graminoids","Forbs", "Woody"))+
  scale_y_continuous(limits=c(0,0.6))+
  ylab("Growth form contribution\n(mean+95%CI)")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.y = element_text(size=18,face="bold"),
        axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      plot.title = element_text(hjust = 0.5,size=18,color="darkblue"))+
  # annotate("text", x=0.62,y=0.6,label="(B)",fontface="bold",size=5)+
  annotate("text", x=1,y=0.23,label="a",size=6) +
  annotate("text", x=2,y=0.25,label="a",size=6) +
  annotate("text", x=3,y=0.24,label="a",size=6) 

ggsave(file.path(graph.wd,"GrowthForms.png"), plot = last_plot(), width=6, height=5, dpi = 600, units="in")

```

(Q2) What characteristics of the native plant community are associated with faster recovery?
(Q3) Is the recovery of the native community affected by environmental conditions, specifically, light and soil water availability?
```{r}
# calculating means, so alphas are comparable
tgfMean<-as.numeric(data %>% select(c("light1","light2","light3")) %>% pivot_longer(1:3) %>% summarise(tgfMean=mean(value)))
smMean<-as.numeric(data %>% select(c("smJL1","smJL2","smJL3")) %>% pivot_longer(1:3) %>% summarise(tgfMean=mean(value)))
RichNMean<-as.numeric(data %>% select(c("RichNJN1","RichNJN2","RichNJN3")) %>% pivot_longer(1:3) %>% summarise(tgfMean=mean(value)))
TraitMean<-as.numeric(data %>% select(c("SLA61","SLA62","SLA63")) %>% pivot_longer(1:3) %>% summarise(tgfMean=mean(value)))

ALPHAmodelpar <- as.data.frame(parameters0[2]) %>% 
  rownames_to_column(., var = "parameter") %>% 
  filter(grepl("^alpha",parameter)) %>% 
  mutate(ID=row_number(),
         `quantiles.50.`=ifelse(ID==1,`quantiles.50.`*mean(RichNMean),
                                ifelse(ID==4,`quantiles.50.`*mean(smMean),
                                       ifelse(ID==3,`quantiles.50.`*mean(tgfMean),`quantiles.50.`*mean(TraitMean)))),
         `quantiles.2.5.`=ifelse(ID==1,`quantiles.2.5.`*mean(RichNMean),
                                 ifelse(ID==4,`quantiles.2.5.`*mean(smMean),
                                        ifelse(ID==3,`quantiles.2.5.`*mean(tgfMean),`quantiles.2.5.`*mean(TraitMean)))),
         `quantiles.97.5.`=ifelse(ID==1,`quantiles.97.5.`*mean(RichNMean),
                                  ifelse(ID==4,`quantiles.97.5.`*mean(smMean),
                                         ifelse(ID==3,`quantiles.97.5.`*mean(tgfMean),`quantiles.97.5.`*mean(TraitMean)))),
         Filling=ifelse(`quantiles.2.5.`<0&`quantiles.97.5.`<0|
                                                              `quantiles.2.5.`>0&`quantiles.97.5.`>0,
                                                            "black","white"),
         ID=ifelse(ID==3,2.9,ifelse(ID==2,1.95,ID)))
ALPHAmodelpar %>% 
  ggplot(aes(y=`quantiles.50.`,x=ID))+
  geom_errorbar(aes(ymin=`quantiles.2.5.`, ymax=`quantiles.97.5.`),width=.1, linewidth=1.3, color="black")+
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=21,size=8,fill=ALPHAmodelpar$Filling)+
  scale_x_continuous(breaks=c(1,1.95,2.9,4),
                      labels=c("Native richness", expression("CWM"[SLA]),"Light availability","Soil moisture"),
                     limits=c(0.6,4.25))+
  ylab("Standardized coefficients\n(mean+95%CI)")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 18),
        axis.title = element_text(size = 20, face="bold"),
        axis.title.x = element_blank(),
        legend.margin=margin(t=-10),
        legend.title = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank()) +
  annotate("text", x=0.6,y=20,label="(A)",fontface="bold",size=6)
  

ggsave(file.path(graph.wd,"CovariatesEffects.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 600, units="in")
```

Vapor Pressure Deficit (VPD)
```{r}
VPD %>% 
  ggplot(aes(x=ID,y=VPD_cumsum,color=as.factor(year))) + 
  geom_line(linewidth=1.5)+
  ylab("Cumulative VPD\n(kPa, based on daily average values)")+
  # labs(y=expression(atop(bold("Cumulative VPD"),atop("(kPa, based on daily average values)")))) +
  scale_color_manual(values=mocolor)+
  scale_y_continuous(n.breaks = 5)+
  scale_x_continuous(breaks=c(1,32,61,93),
                     labels=c("May","June","July","August"),
                     expand=c(0,0.3))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.y = element_text(size = 20,face="bold"),
        axis.title.x = element_blank(),
        legend.margin=margin(-4,3,2,1),
        legend.position = c(0.11,0.8),
        legend.text = element_text(size=15),
        legend.title = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"))+
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
  annotate("text", x=18,y=65,label="(B) Forest 1",fontface="bold",size=6)
ggsave(file.path(graph.wd,"CumVPD_oneSite.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 600, units="in")
```

Which seeded species established the most? What are their characteristics?
```{r}
#rectangles
anno <- data.frame(x1min=c(14.64440-(2*4.957329),26.17733-(2*14.032200),
                           15.83125-(2*2.296895),24.62900-(2*7.969518)), 
                   x1max=c(14.64440+(2*4.957329),26.17733+(2*14.032200),
                           15.83125+(2*2.296895),24.62900+(2*7.969518)),
                   y1min=c(Inf,Inf,Inf,Inf),
                   y1max=c(-Inf,-Inf,-Inf,-Inf),
                   Presence=c(0,1,0,1),
                   MixType=c("(A) Forbs added","(A) Forbs added","(B) Grasses added","(B) Grasses added"))
anno2 <- data.frame(x1=c(45,45),
                    x2=c(45,45),
                    y1=c(0.031,0.031),
                    y2=c(0.031,0.031),
                    MixType=c("(A) Forbs added","(B) Grasses added"),
                    colorseg=c("white","black"))

dev.off()

traitsDiversity %>% 
  ggplot(aes(x = SLA_mm2.mg,y = after_stat(density))) +
  geom_density(color = "black", linewidth = 1.5) +
  geom_errorbarh(aes(y = 0.002,xmin=SLAmean-(2*SLAsd), xmax=SLAmean+(2*SLAsd),color=as.factor(Presence)), traitsSeedMix,height=.003, linewidth=1.3,alpha=0.7)+
  geom_point(aes(y = 0.002,x=SLAmean,color=as.factor(Presence)), traitsSeedMix, shape=19,size=5)+
  geom_rect(data=anno,aes(x = NULL,y = NULL,xmin=x1min,xmax=x1max,ymin=y1min,ymax=y1max,fill=as.factor(Presence)),alpha=0.2)+
  scale_color_manual(values=c(sitecolor[1],sitecolor[2], "black", "white"),labels=c("Failed to\nestablish","Established","Occurring\nnatives",""))+
  scale_fill_manual(values=sitecolor2)+
    geom_segment(data=anno2,aes(x=x1,y=y1,xend=x2,yend=y2,color=colorseg))+
  facet_grid(~MixType)+ 
  xlab(expression(bold("SLA of native species (mm"^2*"/mg)")))+
  ylab(expression(bold("Density")))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
         axis.text = element_text(size = 18),
        axis.title.y = element_text(size = 20,face="bold"),
        axis.title.x = element_text(size = 17,face="bold"),
        strip.text.x = element_text(size = 16,face="bold",hjust=0),
        strip.background = element_blank(),
        legend.title = element_blank(),
        legend.position = c(0.89,0.8),
        legend.text = element_text(size=15),
        legend.background = element_blank(),
        legend.margin=margin(-4,1,1,1),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      axis.text.x = element_text(angle = 0),
        panel.spacing = unit(1, "lines"))+
  guides(fill="none",color=guide_legend(override.aes = list(linetype=c(0,0,1,0),shape=c(16,16,NA,NA)))) 
ggsave(file.path(graph.wd,"TraitsSeededSppv2.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 600, units="in")

traitsSeedMix %>% 
  lm(frequency ~ SLA_mm2.mg, data = .) %>% 
  broom::tidy(conf.int = TRUE)

traitsSeedMix %>% 
  filter(!is.na(frequency)) %>%
  mutate(MixType=ifelse(MixType=="(A) Forbs added","Forb","Grass")) %>% 
  ggplot(aes(x=SLA_mm2.mg,y=frequency,shape=MixType,color=MixType)) +
  geom_point(size=7,alpha=0.7)+
  scale_x_continuous(limits=c(10,55))+
  scale_y_continuous(n.breaks = 7)+
  scale_color_manual(values=c("#737373","black"))+
  ylab("Number of occurences")+
  xlab(expression(bold("SLA of established seeded natives (mm"^2*"/mg)"))) +
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(size = 17,face="bold",hjust=0.5),
        axis.title.y = element_text(size = 20,face="bold"),
        legend.title = element_blank(),
        legend.position = c(0.1,0.85),
        legend.text = element_text(size=15),
        legend.background = element_blank(),
        legend.margin=margin(-4,2,2,2),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black"),
      axis.text.x = element_text(angle = 0)) + 
  guides(shape = guide_legend(override.aes = list(size=5,alpha=1)))+
  annotate("text", x=10,y=142,label="(C)",fontface="bold",size=6)
ggsave(file.path(graph.wd,"TraitsSeededSppFreq.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 600, units="in")
```

## Appendix
Residual plots
```{r residuals}
#res ~ location
p1<-ggplot(data = predModel, aes(y = residuals, x = site)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=19,size=5,alpha=0.2) +
  ylab("Residuals")+ 
  scale_x_discrete(labels=c("Forest 1","Forest 2","Forest 3","Forest 4"))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 13,angle=35),
        axis.title = element_text(size = 17),
                axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      panel.spacing = unit(1, "lines"))+
  annotate("text", x=0.6,y=55,label="(A)",fontface="bold",size=4)

p2<-ggplot(data = predModel, aes(y = residuals, x = forest_stand)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=19,size=5,alpha=0.2) +
  ylab("Residuals")+ 
  scale_x_discrete(labels=c("1A","1B","1C",
                            "2A","2B","2C","2D",
                            "3A","3B",
                            "4A","4B","4C"))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 12,angle=35),
        axis.title = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      panel.spacing = unit(1, "lines"))+
  annotate("text", x=1,y=55,label="(B)",fontface="bold",size=4)

#year
p3<-ggplot(data = predModel, aes(x = as.factor(year), y = residuals)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=19,size=5,alpha=0.2) +
  ylab("Residuals")+ 
  xlab("  ")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 17),
        axis.title.x = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      panel.spacing = unit(1, "lines"))+
  annotate("text", x=0.55,y=55,label="(C)",fontface="bold",size=4)

#cover ias
p4<-ggplot(data = predModel, aes(x = CoverI2019, y = residuals)) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=19,size=5,alpha=0.2) +
  ylab("Residuals")+ 
  xlab("Invasive species cover in 2019 (%)")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 14),
        axis.text.x = element_text(size = 13),
        axis.title.x = element_text(size = 13),
        axis.title.y = element_blank(),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      plot.title = element_text(hjust = 0.5,size=18,color="darkblue"),
      panel.spacing = unit(1, "lines"))+
  annotate("text", x=5,y=55,label="(D)",fontface="bold",size=4)

# overall figure
p1 <- ggplotGrob(p1)
p2 <- ggplotGrob(p2)
p3 <- ggplotGrob(p3)
p4 <- ggplotGrob(p4)
#showing them all
png(file.path(graph.wd, "Residuals.png"),width=7, height=7, res=300,units="in")
grid::grid.draw(cbind(rbind(p1,p3), rbind(p2,p4)))
grid::grid.draw(cbind(rbind(p1,p3), rbind(p2,p4)))
dev.off()
```

Dominant invader
```{r residuals}
ggplot(data = predModel, aes(y = residuals, x = reorder(sppName,desc(sppName)),color=site)) +
  coord_flip()+
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(shape=19,size=5,alpha=0.5) +
  ylab("Residuals")+ 
scale_color_manual(values = sitecolor,labels=c("Forest 1","Forest 2","Forest 3","Forest 4"))+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 13,face = 'italic'),
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 17),
        axis.title.y = element_blank(),
        strip.text.x = element_text(size = 15),
        axis.line = element_line(colour = "black"),
      panel.border = element_blank(),
      plot.title = element_text(hjust = 0.5,size=18,color="darkblue"),
      panel.spacing = unit(1, "lines"),
      legend.position = "bottom",
      legend.margin=margin(t=-10),
      legend.title=element_blank())+   
  guides(color = guide_legend(override.aes = list(alpha=100,size=5)))
ggsave(file.path(graph.wd,"ResidualsIAS.png"), plot = last_plot(), width=7, height=6, dpi = 600, units="in")
```

Exploring effect of native richness
```{r}
RichnessCover <- richnessN %>% 
  group_by(year, plot) %>% 
  mutate(RichnessPlot_N = mean(Richness_N),
         ES_FG_O = (PctCover_100_N[subplot==3]-PctCover_100_N[subplot==1])/(PctCover_100_N[subplot==3]+PctCover_100_N[subplot==1]/2),
         ES_F_O = (PctCover_100_N[subplot==4]-PctCover_100_N[subplot==1])/(PctCover_100_N[subplot==4]+PctCover_100_N[subplot==1]/2),
         ES_FG_M = (PctCover_100_N[subplot==9]-PctCover_100_N[subplot==7])/(PctCover_100_N[subplot==9]+PctCover_100_N[subplot==7]/2),
         ES_F_M = (PctCover_100_N[subplot==10]-PctCover_100_N[subplot==7])/(PctCover_100_N[subplot==10]+PctCover_100_N[subplot==7]/2)) %>% 
  ungroup() %>% 
  pivot_longer(8:11,values_to = "ES",names_to = "Treatment") %>%
  dplyr::select(-subplot,-PctCover_100_N,-Richness_N) %>% 
  mutate(removal=ifelse(grepl("M$",Treatment),"Multi-year removal","One-time removal"),
         seedMix=ifelse(grepl("^ES_FG_",Treatment),"Forbs and grasses added","Forbs added")) %>% 
  distinct() 

RichnessCover$removal <- factor(RichnessCover$removal, levels = c("One-time removal","Multi-year removal"))
RichnessCover$seedMix <- factor(RichnessCover$seedMix, levels = c("Forbs and grasses added","Forbs added"))

RClm <- RichnessCover %>% group_by(removal,seedMix) %>%
   do(fitHour = lm(ES ~ RichnessPlot_N, data = .))

anno <- data.frame(x1=c(2,2,2,2),
                   x2=c(8,8,8,8),
                   y1=c(1.3,1.3,1.3,1.3),
                   removal=c("One-time removal","One-time removal","Multi-year removal","Multi-year removal"),
                   seedMix=c("Forbs and grasses added","Forbs added","Forbs and grasses added","Forbs added"),
                   textAnno=c("(A)","(B)","(C)","(D)"),
                   textAnno2=c(paste0("Slope p-value = ",round(summary(RClm[[3]][[1]])[["coefficients"]][2,4],2)),
                               paste0("Slope p-value = ",round(summary(RClm[[3]][[2]])[["coefficients"]][2,4],2)),
                               paste0("Slope p-value = ",round(summary(RClm[[3]][[3]])[["coefficients"]][2,4],2)),
                               paste0("Slope p-value = ",round(summary(RClm[[3]][[4]])[["coefficients"]][2,4],2))))

anno$removal <- factor(anno$removal, levels = c("One-time removal","Multi-year removal"))
anno$seedMix <- factor(anno$seedMix, levels = c("Forbs and grasses added","Forbs added"))

glimpse(RichnessCover)

RichnessCover %>% 
  ggplot(aes(y=ES,x=RichnessPlot_N)) + 
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(size=7,alpha=0.7,aes(color=as.factor(year))) +
  stat_smooth(method="lm",color="black") + 
  # ggpubr::stat_regline_equation(label.y = 0.8, label.x = 10)+
  facet_grid(removal~seedMix)+
  geom_text(data=anno,aes(x=x1,y=y1,label=textAnno),size=5,fontface="bold")+
  geom_text(data=anno,aes(x=x2,y=y1,label=textAnno2),size=4,fontface="italic")+
  scale_color_manual(values = mocolor)+
  xlab(expression(bold("Native richness (plot mean)")))+
  labs(y=expression(atop(bold("Effect size of native cover"),atop("(Seeds added - Control).|Average|"^-1*"")))) +
  theme_classic()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 15.5),
        axis.title = element_text(size = 18),
        axis.title.y = element_text(vjust = -4),
        strip.text = element_text(size = 15, face="bold"),
        strip.background = element_blank(),
        legend.margin=margin(-4,3,2,1),
        legend.position = c(0.9,0.6),
        legend.text = element_text(size=14),
        legend.background = element_blank(),
        legend.title = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.spacing = unit(1, "lines"))+
  guides(color=guide_legend(override.aes = list(size=4))) 

ggsave(file.path(graph.wd,"ESCoverVsRich.png"), plot = last_plot(), width=6.8, height=8, dpi = 600, units="in")

```

Native cover x year x treatment
```{r}
NatCover2 <- predModel %>%
  mutate(site=ifelse(site=="arb", "Forest 1",
                     ifelse(site=="esgr","Forest 2",
                            ifelse(site=="mbg", "Forest 3","Forest 4")))) %>% 
  group_by(subplotOriginal,year,site) %>% 
  summarise(MeanCov=mean(value),
            CICovLower=MeanCov-(2*sd(value)),
            CICovUpper=MeanCov+(2*sd(value))) %>% 
  ungroup() %>% 
    rename(subplot=subplotOriginal) %>% 
  arrange(subplot) %>% 
  mutate(Treatment=ifelse(subplot<7,1,2),
         Treatment=ifelse(year==2021,Treatment+2,
                          ifelse(year==2022,Treatment+4,Treatment))) %>% 
  arrange(year) %>% 
  mutate(guide = rep(1:3, each=4, times=6)) %>% 
  arrange(year,guide) %>% 
  mutate(index=sort(rep(seq(1,length=18,by=0.3),times=4)),
         index=ifelse(year==2021,index+0.2,
                      ifelse(year==2022,index+0.4,index)),
         type=ifelse(subplot==3|subplot==9,"Forbs and\ngrasses added",
                     ifelse(subplot==4|subplot==10,"Forbs added","Control")),
         type2=ifelse(subplot==3|subplot==9,24,
                     ifelse(subplot==4|subplot==10,22,21)),
         Filling=ifelse(CICovLower<0&CICovUpper<0|CICovLower>0&CICovUpper>0,
                                                            "black","white"),
         removal=ifelse(subplot<5,"One-time removal","Multi-year removal")) %>% 
  group_by(Treatment,year, Filling) %>% 
  mutate(Filling2=cur_group_id())

names(mocolor2) <- c("1","2","3","4","5","6")

NatCover2$type <- factor(NatCover2$type, levels = c("Control","Forbs and\ngrasses added","Forbs added"))


NatCover2 %>% 
  ggplot(aes(y=MeanCov,x=index,shape=as.factor(type),color=as.factor(Treatment),fill=as.factor(Filling2)))+  
  geom_errorbar(aes(ymin=CICovLower, ymax=CICovUpper),
                width=.1, linewidth=1.3)+
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_point(size=6,stroke=2,shape=NatCover2$type2)+
  scale_y_continuous(n.breaks=8,limits=c(NA,170))+
  scale_x_continuous(breaks=c(1.85,3.85,5.85),
                     labels=c("2020","2021","2022"),
                     expand=c(0,0.3))+
  scale_color_manual(values=mocolor2,
                     breaks=c("1","2"),
                     labels=c("One-time removal","Multi-year removal"),
                     guide=guide_legend(order = 1,override.aes = list(shape=16,size=4,linetype = c(0,0),
                                                            color="black",alpha=c(0.2,1),
                                                            margin(-4,0,0,0))))+
  scale_fill_manual(values=c("#DD9EFE","white","#AA0DFE","white","#ADCDFE","white",
                             "#3283FE","white","#CEC19E","white","#85660D","white"),
                    labels=c("Control","Forbs and\ngrasses added",
                             "Forbs added","","","","","","","","",""))+
  ylab("Native cover (%, mean+95%CI)")+
  facet_wrap(~site)+
  theme_classic()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 20),
        axis.text.x = element_text(size=22,color = c(mocolor2[2], mocolor2[4], mocolor2[6])),
        axis.title = element_text(size = 20, face="bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(vjust = 1),
        strip.text.x = element_text(size = 16,face="bold",hjust=0),
        strip.background = element_blank(),
        legend.margin=margin(-4,3,2,1),
        legend.position = c(0.87,0.35),
        legend.text = element_text(size=13),
        legend.background = element_blank(),
        legend.title = element_blank(),
        axis.line = element_line(colour = "black"),
        panel.border = element_blank(),
        panel.spacing = unit(1, "lines")) +
  guides(fill=guide_legend(order = 2,override.aes = list(shape=c(1,2,0,NA,NA,NA,NA,NA,NA,NA,NA,NA),
                                                         size=3.5,
                                                            linetype = c(0,0,0),
                                                            color="black",margin(-4,0,0,0))))

ggsave(file.path(graph.wd,"NatCoverVsYear.png"), plot = last_plot(), width=8.5, height=10, dpi = 600, units="in")
```

What invasive plants were present and most prevalent within each forest type?
```{r}
glimpse(dominantI)
dominantI_subplot <- dominantI %>% 
  left_join(distinct(select(data, c("forest_stand","plot")))) %>% 
  mutate(forest = str_sub(forest_stand, 1, -2)) %>% 
  group_by(forest) %>% 
  add_count(sppName) %>% 
  distinct(sppName,n) %>% 
  ungroup() %>% 
  arrange(forest,sppName) 

dominantI_subplot %>% 
  group_by(forest) %>% 
  slice_max(n) %>%
  ungroup()

dominantI_subplot_summary <- dominantI_subplot %>% 
  group_by(forest) %>% 
  summarize(sppName = str_c(sppName, collapse = "; ")) %>% 
  ungroup()
```

What was the raw change in cover after invasive species removal in response to the various treatments?
```{r}
rawDiff<- data %>%
  mutate(forest = str_to_upper(str_sub(forest_stand, 1, -2))) %>% 
  mutate(raw_diff = Cover3-Cover2019,
         guide = rep(1:3, times=72),
         index = rep(c(1.0,1.9,2.8, 1.3, 2.2, 3.1), times=36),
         Treatment = rep(c(1,3,5, 2, 4, 6), times=36),
         type=ifelse(subplotOriginal==3|subplotOriginal==9,"Forbs and\ngrasses added",ifelse(subplotOriginal==4|subplotOriginal==10,"Forbs added","Control")),
         removal=ifelse(subplotOriginal<5,"One-time removal","Multi-year removal")) %>% 
  arrange(guide)

mocolor2 <- c("gray","black","gray","black","gray","black")
names(mocolor2) <- c("1","2","3","4","5","6")


rawDiff %>% 
  ggplot(aes(x=index,y=raw_diff,color=as.factor(Treatment),shape=as.factor(type))) +
  geom_hline(yintercept=0, color="black", linetype="dashed", alpha=.5, linewidth=1)+
  geom_boxplot()+
  geom_jitter(size=3, alpha=0.5, stroke=2, width = 0.05, height = 0) +
    labs(y = "Difference in native cover\n2023-2019 (%)") + 
  scale_x_continuous(breaks=c(1.15,2.05,2.95),
                     labels=c("Control","Forbs and\ngrasses added","Forbs added"),
                     expand=c(0,0.3))+
  theme_classic()+
  scale_color_manual(values=mocolor2, # Colors for all index values
                     breaks=c("1","2"),
                     labels=c("One-time removal","Multi-year removal"),
                     guide=guide_legend(override.aes = list(shape=16,size=4,linetype = c(0,0),
                                                            color="black",alpha=c(0.2,1))))+
    theme(axis.text = element_text(size = 18),
        axis.title.y = element_text(size = 20,face="bold"),
        axis.title.x = element_blank(),
        legend.position = c(0.15,0.1),
        axis.line = element_line(colour = "black"),
        legend.title = element_blank())+
  guides(shape="none")

ggsave(file.path(graph.wd,"RawDiffNatCov.png"), plot = last_plot(), width=6.8, height=5.2, dpi = 600, units="in")

#statistics
rawDiff %>%
  group_by(subplot,type,removal) %>%
  summarise(
    mean_raw_diff = mean(raw_diff, na.rm = TRUE),
    median_raw_diff = median(raw_diff, na.rm = TRUE),
    sd_raw_diff = sd(raw_diff, na.rm = TRUE),
    min_raw_diff = min(raw_diff, na.rm = TRUE),
    max_raw_diff = max(raw_diff, na.rm = TRUE),
    n = n()  # Number of observations
  )

summary(aov(raw_diff ~ removal + type + removal:type, data = rawDiff))

```

Quantifying effects
```{r}
# Forbs added multi-year removal/forbs added one-time removal
round((Stats[Stats$ParameterID == "B[6,2]", "statistics.Mean"]/Stats[Stats$ParameterID == "B[3,2]", "statistics.Mean"]-1)*100,2)

# 2021 rates were higher than the other two years
Stats %>% 
  filter(grepl("^B", ParameterID)) %>% 
  mutate(year = rep(c(2020,2021,2022), each=6),
         treatment = rep(1:6, times=3)) %>% 
  select(treatment,year,statistics.Mean) %>% 
  pivot_wider(names_from = year, values_from = statistics.Mean) %>% 
  mutate(pct2120 = round((`2021`/`2020`-1)*100,2),
         pct2122 = round((`2021`/`2022`-1)*100,2)) %>% 
  select(pct2120,pct2122) %>% 
  pivot_longer(cols = pct2120:pct2122,names_to = "id",values_to = "pct") %>% 
  group_by(id) %>% 
  summarise(pct = mean(pct))

# Native cover and richer forests
NatCover2 %>% 
  ungroup() %>% 
  mutate(removal=ifelse(subplot<5,"One-time","Multi-year"),
         type=ifelse(subplot==3|subplot==9,"Forbs and\ngrasses added",
                     ifelse(subplot==4|subplot==10,"Forbs added","Control"))) %>% 
  select(type,removal, year, site, MeanCov) %>% 
  pivot_wider(names_from = c(removal), values_from = MeanCov) %>% 
  mutate(pctChange = round((`Multi-year`/`One-time`-1)*100,2)) %>% 
  select(-type,-`One-time`,-`Multi-year`) %>% 
  group_by(site, year) %>% 
  summarise(pctChange = mean(pctChange)) %>% 
  pivot_wider(names_from = c(year), values_from = pctChange) %>% 
  mutate(times_higher_2120 = round(`2021`/`2020`,0),
         times_higher_2220 = round(`2022`/`2020`,0))

# Co-variate effects
Stats %>% 
  filter(grepl("^alpha", ParameterID)) %>% 
  mutate(across(-ParameterID, round, digits = 2))

# established vs not established
traitsSeedMix %>% 
  select(MixType, Presence, SLAmean) %>% 
  distinct() %>% 
  pivot_wider(names_from = c(Presence), values_from = SLAmean) %>% 
  mutate(times_higher = `1`/`0`,
         pct_higher = round((`1`/`0`-1)*100),0)
```